DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-02-02 paper/main.tex
revision
Minor changes to paper
8 -> 9 layers.

%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
\author[1]{Michael G. Bechtel}
\author[2]{Elise McEllhiney}
\author[3]{Minje Kim}
\author[4]{Heechul Yun}
\affil[1]{University of Kansas, Lawrence, United States\\
  \texttt{mbechtel@ku.edu}}
\affil[2]{University of Kansas, Lawrence, United States\\
  \texttt{e908m429@ku.edu}}
\affil[3]{Indiana University, Bloomington, United States\\
  \texttt{minje@indiana.edu}}
\affil[4]{University of Kansas, Lawrence, United States\\
  \texttt{heechul.yun@ku.edu}}
\authorrunning{M.\,G. Bechtel, E. McEllhiney, H. Yun} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'

\Copyright{Michael G. Bechtel, Elise McEllhiney, Heechul Yun}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

@@

\maketitle
\thispagestyle{empty}
\begin{abstract}
We present DeepPicar, a low-cost deep neural network (DNN) based
autonomous car platform. DeepPicar is a small scale
replication of a real self-driving car called Dave-2 by NVIDIA, which
drove on public roads using a deep convolutional neural network (CNN), 
that takes images from a front-facing camera as input and produces
car steering angles as output. DeepPicar uses almost the exact same 
network architecture---8 layers, 27 million connections and 250K
parameters---and can be trained to drive itself, in real-time, using a
web camera and a modest Raspberry Pi 3 quad-core platform.
Using DeepPicar, we analyze the Pi 3's computing capabilities to 
support end-to-end deep learning based real-time control of autonomous
vehicles. We also systematically compare other contemporary embedded
computing platforms using the DeepPicar's CNN based real-time control
software as a workload. 
We find all tested platforms, including the Pi 3, are capable of
supporting deep-learning based real-time control, from 20 Hz up to 100
Hz depending on hardware platform. 
However, shared resource contention remains an
important issue that must be considered in applying deep-learning
models on shared memory based embedded computing platforms.
\end{abstract}

%-------------------------------------------------------------------------
\input{intro}
\input{background}
\input{overview}
\input{evaluation}
\input{comparison}
\input{related}
\input{conclusion}
%-------------------------------------------------------------------------
%\bibliographystyle{abbrv}
\bibliography{reference}

\appendix
\input{appendix}
\end{document}

>>>>

%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
\author[1]{Michael G. Bechtel}
\author[2]{Elise McEllhiney}
\author[3]{Minje Kim}
\author[4]{Heechul Yun}
\affil[1]{University of Kansas, Lawrence, United States\\
  \texttt{mbechtel@ku.edu}}
\affil[2]{University of Kansas, Lawrence, United States\\
  \texttt{e908m429@ku.edu}}
\affil[3]{Indiana University, Bloomington, United States\\
  \texttt{minje@indiana.edu}}
\affil[4]{University of Kansas, Lawrence, United States\\
  \texttt{heechul.yun@ku.edu}}
\authorrunning{M.\,G. Bechtel, E. McEllhiney, M. Kim, H. Yun} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'

\Copyright{Michael G. Bechtel, Elise McEllhiney, Minje Kim, Heechul Yun}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

@@

\maketitle
\thispagestyle{empty}
\begin{abstract}
We present DeepPicar, a low-cost deep neural network (DNN) based
autonomous car platform. DeepPicar is a small scale
replication of a real self-driving car called Dave-2 by NVIDIA, which
drove on public roads using a deep convolutional neural network (CNN), 
that takes images from a front-facing camera as input and produces
car steering angles as output. DeepPicar uses almost the exact same 
network architecture---9 layers, 27 million connections and 250K
parameters---and can be trained to drive itself, in real-time, using a
web camera and a modest Raspberry Pi 3 quad-core platform.
Using DeepPicar, we analyze the Pi 3's computing capabilities to 
support end-to-end deep learning based real-time control of autonomous
vehicles. We also systematically compare other contemporary embedded
computing platforms using the DeepPicar's CNN based real-time control
software as a workload. 
We find all tested platforms, including the Pi 3, are capable of
supporting deep-learning based real-time control, from 20 Hz up to 100
Hz depending on hardware platform. 
However, shared resource contention remains an
important issue that must be considered in applying deep-learning
models on shared memory based embedded computing platforms.
\end{abstract}

%-------------------------------------------------------------------------
\input{intro}
\input{background}
\input{overview}
\input{evaluation}
\input{comparison}
\input{discussion}
\input{related}
\input{conclusion}
%-------------------------------------------------------------------------
%\bibliographystyle{abbrv}
\bibliography{reference}
%% \appendix
\end{document}

### 2018-02-02 paper/overview.tex
8 -> 9 layers.

\begin{figure}[h]
  \centering
  \includegraphics[width=.55\textwidth]{figs/DeepPicar_platform}
  \caption{DeepPicar platform.}
  \label{fig:overview}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.65\textwidth]{figs/architecture}
  \caption{DeepPicar's neural network architecture: 8 layers (5
    convolutional, 3 fully-connected layers), 27 million connections,
    250K parameters. The CNN architecture is identical to the one 
	used in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
  \label{fig:architecture}
\end{figure}

For the neural network architecture, we implement
NVIDIA DAVE-2's convolutional neural network (CNN) using a TensorFlow
based CNN model in ~\cite{deeptesla}. Note, however, that the CNN model
in~\cite{deeptesla} was considerably bigger than NVIDIA's CNN model as
it contained an additional fully-connected layer of approximately 1.3M
additional parameters. We eliminated the additional layer to
faithfully recreate the NVIDIA's original CNN model.
%% The main difference is that we do not utilize a normalization layer, and 
%% instead initialize the weights using a Xavier initialization. 
As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
as input and produces a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture
used in this paper, which is comprised of 8 layers, 250K parameters,
and about 27 million connections as in NVIDIA DAVE-2's. Note, however,
that we did not apply the normalization mentioned
in~\cite{Bojarski2016}, as it does not include trainable parameters
and its computational demand with respect to the overall CNN
processing is minimal.

\begin{figure}[h]
  \centering
%% <<<<<<< HEAD
%%   \includegraphics[width=.5\textwidth]{figs/track_new2}
%%   \caption{One of the custom tracks used for training/testing.}
%%   \label{fig:track}
%% =======
  \begin{subfigure}{0.4\textwidth}
    \includegraphics[width=\textwidth]{figs/track_new2}
    \caption{Track 1.}
    \label{fig:track}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.52\textwidth}
    \includegraphics[width=\textwidth]{figs/track-2}
    \caption{Track 2.}
    \label{fig:track2}
  \end{subfigure}
  \caption{Custom tracks used for training/testing}
  \label{fig:tracks}
%% >>>>>>> f3814396d8701264f2439b42fbe8bbb608c41452
\end{figure}

>>>>

\begin{figure}[h]
  \centering
  \includegraphics[width=.65\textwidth]{figs/DeepPicar_platform}
  \caption{DeepPicar platform.}
  \label{fig:overview}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.65\textwidth]{figs/architecture}
  \caption{DeepPicar's neural network architecture: 9 layers (5
    convolutional, 4 fully-connected layers), 27 million connections,
    250K parameters. The CNN architecture is identical to the one 
	used in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
  \label{fig:architecture}
\end{figure}

For the neural network architecture, we implement
NVIDIA DAVE-2's convolutional neural network (CNN) using a TensorFlow
based CNN model in ~\cite{deeptesla}. Note, however, that the CNN model
in~\cite{deeptesla} was considerably bigger than NVIDIA's CNN model as
it contained an additional fully-connected layer of approximately 1.3M
additional parameters. We eliminated the additional layer to
faithfully recreate the NVIDIA's original CNN model.
%% The main difference is that we do not utilize a normalization layer, and 
%% instead initialize the weights using a Xavier initialization. 
As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
as input and produces a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture
used in this paper, which is comprised of 9 layers, 250K parameters,
and about 27 million connections as in NVIDIA DAVE-2's. Note, however,
that we did not apply the normalization mentioned
in~\cite{Bojarski2016}, as it does not include trainable parameters
and its computational demand with respect to the overall CNN
processing is minimal.

\begin{figure}[h]
  \centering
%% <<<<<<< HEAD
%%   \includegraphics[width=.5\textwidth]{figs/track_new2}
%%   \caption{One of the custom tracks used for training/testing.}
%%   \label{fig:track}
%% =======
  \begin{subfigure}{0.6\textwidth}
    \includegraphics[width=\textwidth]{figs/track_new2}
    \caption{Track 1.}
    \label{fig:track}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.7\textwidth}
    \includegraphics[width=\textwidth]{figs/track-2}
    \caption{Track 2.}
    \label{fig:track2}
  \end{subfigure}
  \caption{Custom tracks used for training/testing}
  \label{fig:tracks}
%% >>>>>>> f3814396d8701264f2439b42fbe8bbb608c41452
\end{figure}
